---
title: "US model"
date: "`r Sys.time()`"
author: Claudio Zanettini
categories:
  - covid-19
tags: 
  - covid-19
  - Italy
lastmod: "`r Sys.time()`"
keywords: []
description: ''
output:
  html_document:
    df_print: kable
    toc: true
    theme: journal
---

# US

## Get the data and functions

This is to get the data and clean them up.

```{r get_data3, message=FALSE, warning=FALSE}
# libraries --------------
# devtools::install_github("zeehio/facetscales")
library(facetscales)
library(covid19census)
library(tidyverse)
library(ggcorrplot)
library(GGally)
library(gganimate)
library(ggcorrplot)
library(broom)
library(purrr)
library(knitr)
library(kableExtra)
library(htmlwidgets)
library(plotly)
library(stringr)
theme_set(theme_bw())



logit = function(p){ log(p) - log(1-p) }

# var new to select -------
to_select<- c("date", "county", "state", "fips", "cases", "deaths", "perc_families", 
"perc_family_only_onep", "perc_edu_bachelor", 
"perc_withinternet", "total_pop", "perc_imm65", "total_imm65", 
"total_beds", "ratio_beds", "urban",  
"perc_alzheimer_dementia", "perc_asthma", "perc_atrial_fibrillation", 
"perc_cancer_breast", "perc_cancer_colorectal", "perc_cancer_lung", 
"perc_ch_obstructive_pulm", "perc_chronic_kidney_disease", "perc_depression", 
"perc_diabetes", "perc_heart_failure", "perc_hypertension", "perc_ischemic_heart_disease", "perc_obesity", "perc_osteoporosis", "perc_rheumatoid_arthritis", 
 "perc_stroke", "perc_tobacco_use", 
"median_income", "pm2.5", "summer_temp", "summer_hum", "winter_temp", 
"winter_hum", "age65_over", "median_age", "sex_ratio", "child_dependency", 
"perc_black", "perc_lat", "perc_white", "perc_asian", "perc_island", 
"perc_other", "perc_two_more_races", "total_tests", "days_f0"
)



# 
# # var to select old -------
# to_select <- c(
#   "date",
#   "county",
#   "state",
#   "fips",
#   "cases",
#   "deaths",
#   "perc_families",
#   "perc_familes_18childreen",
#   "perc_married_couples",
# 
#   "perc_family_only_onep",
#   # "perc_married_couples_u18ychildreen",
#   # "perc_families_only_male",
#   # "perc_families_only_male__18ychildreen",
#   # "perc_families_only_female",
#   # "perc_families_only_female__18ychildreen",
#   # "perc_non_families_alone",
#   # "perc_non_families_alone65y",
#   # "perc_non_families_u18y",
#   # "perc_non_families_65y",
#   # "perc_relationship_householder",
#   # "perc_relationship_spouse",
#   # "perc_relationship_child",
#   # "perc_relationship_other_relatives",
#   # "perc_relationship_other_nonrelatives",
#   # "perc_relationship_other_unmaried_part",
#   # "perc_marital_status_male_nevermaried",
#   # "perc_marital_status_male_maried",
#   # "perc_marital_status_male_separated",
#   # "perc_marital_status_male_widowed",
#   # "perc_marital_status_male_divorced",
#   # "perc_marital_status_female_nevermaried",
#   # "perc_marital_status_female_maried",
#   # "perc_marital_status_female_separated",
#   # "perc_marital_status_female_widowed",
#   # "perc_marital_status_female_divorced",
#   "perc_enrolled_preschool",
#   "perc_enrolled_kindergarden",
#   "perc_enrolled_elementary",
#   "perc_enrolled_highschool",
#   "perc_enrolled_college",
#   # "perc_edu_9grade",
#   # "perc_edu_nodiploma",
#   # "perc_edu_highshool",
#   "perc_edu_somecollege",
#   "perc_edu_associate",
#   "perc_edu_bachelor",
#   # "perc_edu_gradprofess",
#   # "perc_edu_highshool_higher",
#   # "perc_edu_batchelor_higher",
#   # "perc_american",
#   # "perc_arab",
#   # "perc_czech",
#   # "perc_danish",
#   # "perc_dutch",
#   # "perc_english",
#   # "perc_french_except_basque",
#   # "perc_french_canadian",
#   # "perc_german",
#   # "perc_greek",
#   # "perc_hungarian",
#   # "perc_irish",
#   # "perc_italian",
#   # "perc_lithuanian",
#   # "perc_norwegian",
#   # "perc_polish",
#   # "perc_portuguese",
#   # "perc_russian",
#   # "perc_scotch_irish",
#   # "perc_scottish",
#   # "perc_slovak",
#   # "perc_subsaharan_african",
#   # "perc_swedish",
#   # "perc_swiss",
#   # "perc_ukrainian",
#   # "perc_welsh",
#   # "perc_west_indian_excluding_hispanic_origin_groups",
#   # "perc_withcomputer",
#   "perc_withinternet",
#   "total_pop",
#   "perc_imm65",
#   "total_imm65",
#   "total_beds",
# 
#   "ratio_beds",
#   "urban",
#   "perc_acute_myocardial_infarction",
#   "perc_alzheimer_dementia",
#   "perc_asthma",
#   "perc_atrial_fibrillation",
#   "perc_cancer_breast",
#   "perc_cancer_colorectal",
#   "perc_cancer_lung",
#   "perc_cancer_all",
#   "perc_ch_obstructive_pulm",
#   "perc_chronic_kidney_disease",
#   "perc_depression",
#   "perc_diabetes",
#   "perc_heart_failure",
#   "perc_hypertension",
#   "perc_ischemic_heart_disease",
#   "perc_obesity",
#   "perc_osteoporosis",
#   "perc_rheumatoid_arthritis",
#   "perc_schizophrenia_psychotic_dis",
#   "perc_stroke",
#   "perc_tobacco_use",
#   # to many NA for these:
#   # "urgent_admission",
#   # "annual_wellness_visit",
#   # "elective_admission",
#   # "emergent_admission",
#   # "other_admission",
#   # perc_pneumococcal_vaccine HIGHLY CORRELATED WITH perc_imm65
#   # "perc_pneumococcal_vaccine",
#   # "perc_poverty",
#   "median_income",
#   "pm2.5",
#   "summer_temp",
#   "summer_hum",
#   "winter_temp",
#   "winter_hum",
#   "age65_over",
#   "median_age",
#   # "median_age_male",
#   # "median_age_female",
#   "sex_ratio",
#   # "age_dependency",
#   # "old_age_dependency",
#   "child_dependency",
#   "perc_black",
#   "perc_lat",
#   "perc_white",
# 
#   "perc_asian",
#   "perc_island",
#   "perc_other",
#   "perc_two_more_races",
#   "total_tests",
#   "days_f0"
# )
# 
# 

# df1   -----------------------------
df1 <- getus_all()

# df2 ------------------------------
df2 <-
  df1 %>%
  # calculate age65_over and perc_black, white and latinos, total tests
  mutate(age65_over = rowSums(dplyr::select(., matches("perc_[6][5-9]|[7|8][0-9]")))) %>%
  mutate(urban = if_else(urban == "Urban", 1, 0)) %>%
  mutate(perc_black = total_black / total_pop * 100) %>%
  mutate(perc_white = total_white / total_pop * 100) %>%
  mutate(perc_lat = total_latino / total_pop * 100) %>%
  mutate(total_tests = positive + negative) %>%

  # calculate day since first case
  mutate(f_date = case_when(cases >= 1 ~ date)) %>%
  group_by(fips) %>%
  mutate(f_date = min(f_date, na.rm = TRUE), days_f0 = as.numeric(date - f_date)) %>%
  ungroup() %>%
  mutate(days_f0 = if_else(is.finite(days_f0), days_f0, NA_real_)) %>%
  
  # newly calculated-----------------------------------  
  # total beds in ratio
  mutate(ratio_beds = total_beds/total_pop) %>% 
  # asian are here
  mutate(perc_asian = total_asian / total_pop * 100) %>%
  mutate(perc_island = total_pacific_islander / total_pop * 100) %>%
  mutate(perc_native = total_native / total_pop * 100) %>%
  mutate(perc_other = total_other_race / total_pop * 100) %>%
  mutate(perc_two_more_races = total_two_more_races / total_pop * 100) %>%
  
  
  # family with one parent together
  mutate(perc_family_only_onep = perc_families_only_female + perc_families_only_male) %>% 
  mutate(total_imm65 = age65_over/100 *perc_imm65) %>% 
   
  # select variables ----------
  dplyr::select(!!to_select) %>%

  # perc divided by 100
  mutate_at(vars(starts_with("perc")), function(x) x / 100) %>%
  mutate(age65_over = age65_over / 100) %>%

  # we remove NAs
  na.omit() %>%

  # at least 1 death
  filter(deaths >= 1)


# all_dat -------------
all_dat <- df2 %>%
  # make variables numeric
  mutate_at(vars(
    -date, -county, -state, -fips), 
     as.numeric
  ) %>%
  # rename
  # add XX_ to all the variables excepts those
  rename_at(
    vars(
      -date, -county, -state, -fips, -cases, -deaths,
      -perc_imm65, -cases, -total_pop, -total_imm65
    ),
    ~ paste0("XX_", .)
  ) %>%
  rename_at(vars(perc_imm65, total_imm65), ~ paste0("ZZ_", .)) %>%
  rename_at(vars(deaths), ~ paste0("YY_", .)) %>%
  rename_at(vars(cases), ~ paste0("NC_", .)) %>%
  rename_at(vars(total_pop), ~ paste0("NP_", .)) %>%
  # calculate logitZZ
  mutate(logitZZ_perc_imm65 = logit(ZZ_perc_imm65)) 
  # mutate(logitZZ_total_imm65 = log(ZZ_total_imm65) - log(1 - ZZ_total_imm65))
  

# today_dat ------------------------------
today_dat <- all_dat %>%
  filter(date == max(date))

# functions -----
logit = function(p){ log(p) - log(1-p) }

# functions ----------------------
#' get_PPScore
#'
#' Calculate propensity score and relative glm
#'
#' @param dat dataframe (need to some specific names ecc)
#' @param filt_cases threeshold of number of cases x county to be included in the analysis
#' @param form  formula to use for propensity score, this is a `reformulate` expression 
#' @param offset_f the offset  "log(NC_cases)"
#' @param var_dep dependent variable
#' @param var_int variable of interest
#' @param quant boolean, quantile of propensity score
#' @param include_model include a glm in the datafram
#' @param verbose boolean
#' @param nation variables to reports are different for italy and us. For example Italy doesn't have counties. We need to tell the function so it produced different outputs
#' @return dataframe
#'
get_PPScore <- function(
                        dat = .x, 
                        filt_cases = 1,
                        form = form_ps, 
                        offset_f = "log(NC_cases)",
                        var_dep =  "YY_deaths",
                        var_int = "ZZ_perc_imm65",
                        quant = FALSE,
                        include_model = FALSE,
                        verbose = TRUE,
                        nation = "us"
                        ){

# given a covid dataframe with 1 date, it calculate effect of the indipendent variable using a propensity score
  
  dat <- dat %>%
    filter(dat$NC_cases >= !!filt_cases)
  
  PropScoresLM <- lm(form, data = dat)
  
  if(quant == TRUE){
    dat[, "PP"] <- cut(fitted.values(PropScoresLM), 4, labels = FALSE)
  } else {
    dat[, "PP"] <- fitted.values(PropScoresLM)
  }
  
  form_glm <- reformulate(termlabels = c(var_int, "PP"), response = var_dep)

  outcome_PP <-
    # we add on the fly PP to the dataframe and we
    dat %>%
    glm(form_glm,
        family = quasipoisson(link = "log"),
        offset = eval(parse(text = offset_f)), 
        data = .
    )

  # conf.int = TRUE for confidence intervals
  res <- tidy(outcome_PP,  conf.int = TRUE) %>%
    filter(term == !!var_int) %>%
    mutate(
      deaths = sum(dat$YY_deaths),
      cases = sum(dat$NC_cases),
      formula =  deparse(formula(outcome_PP)),
      PP_form = paste0(deparse(form), collapse = ""),
      offset_f = offset_f,
      quant = FALSE
    )
  
  
  if(nation == "us") {
    res <- 
      res %>% 
      mutate(
      n_counties = nrow(dat),
      urban = sum(dat$XX_urban)
      )
    
  }
  
  if(include_model == TRUE) {
    res <- res %>% 
      mutate(model = list(outcome_PP))
  }
  
 if(verbose == TRUE){
       message(
          cat(paste0(
            "\nDate ",
            first(dat$date2),
            ":\nApplying model ",
            first(res$formula),
            "\noffset: ",
            first(offset_f),
            "\nfilt_cases = ",
            filt_cases,
            "!"
          ))
       )
}
 res
}


#' glm_pp
#'
#' Given a covid-19 dataframe, apply the propensity glm analysis to subset of dates, varying inclusion criteria 
#' per county
#'
#' @param dat dataframe (need to some specific names ecc)
#' @param filt_cases numeric vector of threasholds of cases X county  (default = 1)
#' @param form  formula to use for propensity score, this is a `reformulate` expression 
#' @param offset_f the offset, es "log(NC_cases)"
#' @param var_dep dependent variable, es. "YY_death"
#' @param var_int indipendent variable, es ZZ_perc_imm65"
#' @param data_term if TRUE keeps that dataframe of data used for the analysis
#' @param quant boolean, quantile of propensity score
#' @param include_model include a glm in the datafram
#' @param verbose boolean
#'
#' @return dataframe
#'


glm_pp <- function(dat, 
                   dates_tostudy = dates_tostudy, 
                   filt_cases,
                   offset_f = "log(NC_cases)",
                   var_dep,
                   var_int,
                   data_term = FALSE, 
                   quant = FALSE, 
                   include_model = FALSE,
                   verbose = TRUE,
                   ...
                   ){

# functions to apply glm ppcs model to diffenret dates in dataframe
# dat = covid dataframe 
# dates_tostudy = list of dates to perform the analysis on                    
# filt_cases =  numeric vector of threasholds of cases X county  (default = 1)
# form = formula to use for propensity score this is a `reformulate` expression for the pp score (form_ps)
# offset_f = the offset of glm function ("log(NC_cases)")
# var_dep =  dependent variable ("YY_death")
# var_int = variable of interest ("ZZ_perc_imm65"),
# quant = propensity score quantile (FALSE)


  suppressWarnings(
  zz_term <-   
  dat %>%
  # filter(YY_deaths > 0) %>% 
  filter(date %in% dates_tostudy) %>%
  mutate(date2 = date) %>% 
  nest(-date) %>%
  # we repeat the date slice X number of case threeshold that we want to study
  slice(rep(1:n(), each = length(filt_cases))) %>%
  mutate(filt_cases_c = rep(filt_cases, nrow(.)/length(filt_cases))) %>%
  
  mutate(fit = 
            map2(
            data,
            filt_cases_c,
            get_PPScore,
            form = form_ps, 
            offset_f = !!offset_f,
            quant = !!quant,
            var_int = !!var_int,
            var_dep = !!var_dep,
            include_model = !!include_model,
            verbose = !!verbose
            )
  ) %>% 
    unnest(fit))
    
  if(data_term == FALSE){
    zz_term_nd <- zz_term %>% 
    select(-data) 
    return(zz_term_nd)
  }else{
    return(zz_term)
    } 
}

ggcovid <- function(dat, x_var, filt_cases = 1, nation = "us") {
  
if(!x_var %in% c("date", "cases")) stop("x_var can be date or cases!")

if (x_var == "date"){
  to_add <- "; cases > 0" 
  aes_x <- "date"
  dat <- dat %>%
    filter(filt_cases_c == !!filt_cases) 
  x_title = NULL
} 

if (x_var == "cases") {
  to_add <- paste0("; date ",  max(dat$date))
  aes_x <- "filt_cases_c"
  dat <- dat %>%
    filter(date == max(date))  
  x_title = "inclusion criteria (minimum cases)"
  
}

if (nation == "it") {
  var_facet <- c("estimate", "deaths", "cases")
} 

  
if (nation == "us") {
  var_facet <-  c("estimate", "number of counties", "urban/counties")
}
  
  
# dependent variable used in title
dep_var <- stringr::word(first(dat$formula), 1)



  dat %>% 
    mutate(`urban/counties` = urban/n_counties) %>% 
    rename(`number of counties` = n_counties) %>% 
    pivot_longer(cols =   !!var_facet,
                 names_to = "var_name") %>%
    mutate(var_name = factor(var_name, levels = !!var_facet)) %>% 
    mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
    mutate(sign = factor(sign, levels = c("p<0.05", "NS"))) %>% 
    # mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%
    
    mutate_at(vars(std.error, conf.low, conf.high), ~if_else(var_name == "estimate", ., 0)) %>% 
    
    ggplot(aes_string(aes_x, "value", color = "sign", group = "term")) +
    geom_line() +
    geom_point() +
    geom_linerange(aes(ymax = conf.high, ymin = conf.low), show.legend = FALSE) +
    facet_grid(vars(var_name), scales = "free") +
    scale_color_manual(values = c("NS" = "#00BFC4", "p<0.05" = "#F8766D"), drop = FALSE) +
    theme(
      # legend.position = "top",
      legend.title = element_blank(),
      legend.box.background = element_rect(colour = "black")
    ) +
    labs(
      title = paste0("Slope of ", first(dat$term), " over time"),
      subtitle = paste0(dep_var, "; offset = ", first(dat$offset_f),
                        ";\\nquantile PP = ", first(dat$quant), to_add),
      x = x_title
    ) +
    expand_limits(y = 0) 

}

```



## Results


### Deaths ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)"

```{r  fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zz_term_deaths_np <- glm_pp(dat = all_dat,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )


zz_term_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()


ggcovid(zz_term_deaths_np, "date")

```


```{r fig.width=4, fig.height=4}
ggcovid(zz_term_deaths_np, "cases")
```


### Deaths ~ perc_imm65 + PP, offset_f = "log(NC_cases)"

```{r  fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zz_term_deaths_nc <- glm_pp(dat = all_dat,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zz_term_deaths_nc %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()


ggcovid(zz_term_deaths_nc, "date")
```

```{r fig.width=4, fig.height=4}
ggcovid(zz_term_deaths_nc, "cases")
```

### Only New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)"

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zzterm_onlyNY_deaths_np  <- 
  
  all_dat %>% 
  filter(state == "New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_onlyNY_deaths_np  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()



ggcovid(zzterm_onlyNY_deaths_np, "date")
```


```{r fig.width=4, fig.height=4}
ggcovid(zzterm_onlyNY_deaths_np, "cases")
```

### Only New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NC_cases)"

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "ZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zzterm_onlyNY_deaths_nc  <- 
  
  all_dat %>% 
  filter(state =="New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_onlyNY_deaths_nc %>% 
    filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()
 



ggcovid(zzterm_onlyNY_deaths_nc, "date")
```


```{r fig.width=4, fig.height=4}
ggcovid(zzterm_onlyNY_deaths_nc, "cases")
```


so the estimate is not exactly the same


```{r}
zzterm_onlyNY_deaths_nc %>% 
  filter(date == max(date)) %>% 
  select(estimate)
```



### No New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)"


```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zzterm_noNY_deaths_np  <- 
  all_dat %>% 
  filter(state !="New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_noNY_deaths_np  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()


ggcovid(zzterm_noNY_deaths_np, "date")
```




```{r fig.width=4, fig.height=4}
ggcovid(zzterm_noNY_deaths_np, "cases")
```

### No New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NC_cases)"

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zzterm_noNY_deaths_nc  <- 
  
  all_dat %>% 
  filter(state != "New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_noNY_deaths_nc  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()



ggcovid(zzterm_noNY_deaths_nc, "date")
```

```{r fig.width=4, fig.height=4}
ggcovid(zzterm_noNY_deaths_nc, "cases")
```




With this we check the actual data to make sure that the function actually took all but NYS. We just take the last day and
we filter per counties > 0

```{r}
no_ny_check  <- 
  all_dat %>% 
  filter(state !="New York") %>% 
      glm_pp(
       dates_tostudy = max(all_dat$date),
       filt_cases = 1,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = TRUE,
       include_model = TRUE,
       verbose = FALSE
       )


data_no_ny_check <- no_ny_check$data[[1]]

unique(data_no_ny_check$state)


no_ny_check$formula

```



### NC_cases ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)


```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zz_term_cases_np <- 
  
  all_dat %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form =  form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = TRUE,
       include_model = TRUE,
       verbose = TRUE
       )

 zz_term_cases_np  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()



ggcovid(zz_term_cases_np, "date")

```

```{r fig.width=4, fig.height=4}
ggcovid(zz_term_cases_np, "cases")
```

### No New York: NC_cases ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


zz_term_noNY_cases_np <- 
  
  all_dat %>% 
    filter(state != "New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = TRUE,
       include_model = TRUE,
       verbose = TRUE
       )

 zz_term_noNY_cases_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()



ggcovid(zz_term_noNY_cases_np, "date")

```


### Positive control: pm2.5



```{r}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds", "XX_pm2.5")]

XX <- c("ZZ_perc_imm65", XX )

form_ps <- reformulate(termlabels = XX, response = "XX_pm2.5")


dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


pm2.5_deaths_np <- 
      
      all_dat %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "XX_pm2.5",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )


pm2.5_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(pm2.5_deaths_np, "date")


```

### Positive control Black


with logit doesnt fit

```{r}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds", "XX_perc_black")]

XX <- c("ZZ_perc_imm65", XX )

form_ps <- reformulate(termlabels = XX, response = "XX_perc_black")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- 1   


black_deaths_np <- 
      all_dat %>% 
      mutate(logitXX_perc_black = logit(XX_perc_black)) %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "XX_perc_black",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )


black_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(black_deaths_np, "date")



```
### Positive control income ????

```{r}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds", "XX_median_income")]

XX <- c("ZZ_perc_imm65", XX )

form_ps <- reformulate(termlabels = XX, response = "XX_median_income")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))   


income_deaths_np <- 
  
  all_dat %>% 
  glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "XX_median_income",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )


income_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(income_deaths_np, "date")

```


## Italy

### Deaths


```{r}
dat_it <- getit_all() %>% 
  mutate_at(vars(imm65, chronic_diab, chronic_bronc, chronic_hyp), function(x) x / 100) %>% 
  mutate(logitZZ_perc_imm65 = logit(imm65))

today_dat_it <-  dat_it %>% 
  filter(date == max(date))


dat_it_cl <- dat_it %>% 
  rename_at(vars(imm65), ~ paste0("ZZ_perc_", .)) %>% 
  rename_at(vars(pop_km2, lat, chronic_diab, chronic_bronc, chronic_hyp), ~paste0("XX_", .)) %>% 
  rename_at(vars(total_cases),  ~("NC_cases")) %>% 
  rename_at(vars(pop_tot), ~("NP_total_pop")) %>% 
  rename_at(vars(deaths), ~("YY_deaths"))



XX <- grep("XX", names(dat_it_cl), value = TRUE)
# XX <- XX[!XX %in% c("XX_total_beds") ]

dates_tostudy <- sort(unique(dat_it_cl$date))[52:length(unique(dat_it_cl$date))]
to_filt_cases  <- 1 

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")


perc_imm65_deahts_pop <- glm_pp(dat = dat_it_cl,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE,
       nation = "it"
       )



ggcovid(perc_imm65_deahts_pop, "date", nation = "it")
```

```{r}


XX <- grep("XX", names(dat_it_cl), value = TRUE)
# XX <- XX[!XX %in% c("XX_total_beds") ]



dates_tostudy <- sort(unique(dat_it_cl$date))[52:length(unique(dat_it_cl$date))]
to_filt_cases  <- 1 

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")


perc_imm65_deahts_cases <- glm_pp(dat = dat_it_cl,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE,
       nation = "it"
       )



ggcovid(perc_imm65_deahts_cases, "date", nation = "it")
```

### Cases


```{r}
XX <- grep("XX", names(dat_it_cl), value = TRUE)
# XX <- XX[!XX %in% c("XX_total_beds") ]


dates_tostudy <- sort(unique(dat_it_cl$date))[52:length(unique(dat_it_cl$date))]
to_filt_cases  <- 1 

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")


perc_imm65_cases_cases <- glm_pp(dat = dat_it_cl,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE,
       nation = "it"
       )



ggcovid(perc_imm65_cases_cases, "date", nation = "it")

```






