---
title: "Select Var"
output:
  html_document:
    df_print: kable
    toc: true
    theme: journal
---

Analysis based on GP example.

# Data transformation

Summary of the data transformation :

  * `all_dat` = all data.
  * `all_dat2` = calculated some percentage and days from first case
  * `all_data3` = selected variables (excluded totals and absolute values) and empty, and last date
  * `all_data4` = percentages multiplied by 100, **filtered death >= 1**

then based on GP analysis,the dataframe is spitted in:

  * `XX` = all independent variables/confounders excluding `perc_imm65`
  * `ZZ` = `perc_imm65`
  * `logitZZ` =  `perc_imm65` predicted probability: log[p/(1 − p)]
  * `YY` = dependent variable `deaths`
  * `NC` = number of cases: `cases`
  * `NP` = population: `total_pop`
  * `PP` = fitted values of propensity score linear model
  

```{r libs, message=FALSE, warning=FALSE}
# devtools::install_github("c1au6i0/covid19census_dev",  auth_token = "8ca5cd2bb556a45f1ffeb2e9b44c90d4fbef6cf2")
library(covid19census)
library(tidyverse)
library(ggcorrplot)
library(caret)
library(GGally)
library(ggcorrplot)
library(ggdendro)
library(broom)
```


Get the data, compute some percentage, and the number of days since the first case in each county.

```{r all_dat2, warning=FALSE, message=FALSE}
all_dat <- getus_all()

all_dat2 <-
  all_dat %>%  
  # get 65 and over
  mutate(age65_over = rowSums(dplyr::select(., matches("perc_[6][5-9]|[7|8][0-9]")))) %>% 
  mutate(urban = if_else(urban == "Urban", 1, 0)) %>% 
  mutate(perc_black = total_nlat_blackaa_alone/total_pop * 100) %>% 
  mutate(perc_white = total_nlat_white_alone/total_pop * 100) %>% 
  mutate(perc_lat = total_lat/total_pop * 100) %>% 
  mutate(total_tests = positive + negative) %>% 
  mutate(f_date = case_when(cases >= 1 ~ date)) %>% 
  group_by(fips) %>% 
  mutate(f_date = min(f_date, na.rm = TRUE), days_f0 = as.numeric(date - f_date)) %>% 
  ungroup() 
```




List of variables to select for analysis. Totals (e.s total respondents for a specific question, age bins), and  variables with too many `NA` are not included.

```{r to_select}
to_select <- c(
  "date",
  "county",
  "state",
  "fips",
  "cases",
  "deaths",
  "perc_families",
  "perc_familes_18childreen",
  "perc_married_couples",
  "perc_married_couples_u18ychildreen",
  "perc_families_only_male",
  "perc_families_only_male__18ychildreen",
  "perc_families_only_female",
  "perc_families_only_female__18ychildreen",
  "perc_non_families_alone",
  "perc_non_families_alone65y",
  "perc_non_families_u18y",
  "perc_non_families_65y",
  "perc_relationship_householder",
  "perc_relationship_spouse",
  "perc_relationship_child",
  "perc_relationship_other_relatives",
  "perc_relationship_other_nonrelatives",
  "perc_relationship_other_unmaried_part",
  "perc_marital_status_male_nevermaried",
  "perc_marital_status_male_maried",
  "perc_marital_status_male_separated",
  "perc_marital_status_male_widowed",
  "perc_marital_status_male_divorced",
  "perc_marital_status_female_nevermaried",
  "perc_marital_status_female_maried",
  "perc_marital_status_female_separated",
  "perc_marital_status_female_widowed",
  "perc_marital_status_female_divorced",
  "perc_enrolled_preschool",
  "perc_enrolled_kindergarden",
  "perc_enrolled_elementary",
  "perc_enrolled_highschool",
  "perc_enrolled_college",
  "perc_edu_9grade",
  "perc_edu_nodiploma",
  "perc_edu_highshool",
  "perc_edu_somecollege",
  "perc_edu_associate",
  "perc_edu_bachelor",
  "perc_edu_gradprofess",
  "perc_edu_highshool_higher",
  "perc_edu_batchelor_higher",
  "perc_american",
  "perc_arab",
  "perc_czech",
  "perc_danish",
  "perc_dutch",
  "perc_english",
  "perc_french_except_basque",
  "perc_french_canadian",
  "perc_german",
  "perc_greek",
  "perc_hungarian",
  "perc_irish",
  "perc_italian",
  "perc_lithuanian",
  "perc_norwegian",
  "perc_polish",
  "perc_portuguese",
  "perc_russian",
  "perc_scotch_irish",
  "perc_scottish",
  "perc_slovak",
  "perc_subsaharan_african",
  "perc_swedish",
  "perc_swiss",
  "perc_ukrainian",
  "perc_welsh",
  "perc_west_indian_excluding_hispanic_origin_groups",
  "perc_withcomputer",
  "perc_withinternet",
  "total_pop",
  "perc_imm65",
  "total_beds",
  "urban",
  "perc_acute_myocardial_infarction",
  "perc_alzheimer_dementia",
  "perc_asthma",
  "perc_atrial_fibrillation",
  "perc_cancer_breast",
  "perc_cancer_colorectal",
  "perc_cancer_lung",
  "perc_cancer_all",
  "perc_ch_obstructive_pulm",
  "perc_chronic_kidney_disease",
  "perc_depression",
  "perc_diabetes",
  "perc_heart_failure",
  "perc_hypertension",
  "perc_ischemic_heart_disease",
  "perc_obesity",
  "perc_osteoporosis",
  "perc_rheumatoid_arthritis",
  "perc_schizophrenia_psychotic_dis",
  "perc_stroke",
  "perc_tobacco_use",
  # to many NA for these:
  # "urgent_admission",
  # "annual_wellness_visit",
  # "elective_admission",
  # "emergent_admission",
  # "other_admission",
  "perc_pneumococcal_vaccine",
  "perc_poverty",
  "median_income",
  "pm2.5",
  "summer_temp",
  "summer_hum",
  "winter_temp",
  "winter_hum",
  "age65_over",
  "median_age",
  "median_age_male", 
  "median_age_female", 
  "sex_ratio",
  "age_dependency",                                   
  "old_age_dependency",                              
  "child_dependency",
  "perc_black",
  "perc_lat",
  "perc_white",
  "total_tests",
   "days_f0"
)

```


We filter the dataframe based on that vector.

```{r all_dat3}
all_dat3 <- all_dat2 %>% 
  select(!!to_select) %>% 
  filter(date == max(date)) %>% 
   mutate(days_f0 = if_else(is.finite(days_f0), days_f0, NA_real_))
```

We check the total `NAs` for each variable.

```{r tot_na}
all_na <- as.data.frame(apply(all_dat3, 2, function(x) sum(is.na(x))))
names(all_na) <- "tot_na"

all_na[, "var"] <- row.names(all_na)
all_na %>% 
  arrange(desc(tot_na))
```

It seems that we miss a lot of values for `total beds` (and admission related variables, that have been excluded).


No variable has near zero variance.

```{r zerovar}
library(caret)
nearZeroVar(all_dat3[8:ncol(all_dat3)])
```


We finally divide by 100 all the variables that were percentage.

** We also need to remove the 0 because we are doing a logist regression*

```{r all_dat4}
all_dat4 <-  all_dat3 %>% 
  mutate_at(vars(starts_with("perc")), function(x) x / 100) %>% 
  mutate(age65_over = age65_over / 100) %>% 
  ungroup() %>% 
  na.omit() %>% 
  filter(deaths >= 1)
```

# Dimensionality reduction

## CORRELATIONS

```{r XX, warning=FALSE, message=FALSE}
XX <- all_dat4[, !names(all_dat4) %in% c(
    "date",
    "county",
    "fips",
    "state",
    "cases",
    "deaths",
    "total_pop",
    "perc_imm65"
  )] %>% 
  mutate_all(as.numeric)

```


```{r fam, fig.height=12, fig.width=12}
XX %>% 
  dplyr::select(
    "perc_families", "perc_familes_18childreen", "perc_married_couples", 
"perc_married_couples_u18ychildreen", "perc_families_only_male", 
"perc_families_only_male__18ychildreen", "perc_families_only_female", 
"perc_families_only_female__18ychildreen", "perc_non_families_alone", 
"perc_non_families_alone65y", "perc_non_families_u18y", "perc_non_families_65y", 
"perc_relationship_householder", "perc_relationship_spouse", 
"perc_relationship_child", "perc_relationship_other_relatives", 
"perc_relationship_other_nonrelatives", "perc_relationship_other_unmaried_part", 
"perc_marital_status_male_nevermaried", "perc_marital_status_male_maried", 
"perc_marital_status_male_separated", "perc_marital_status_male_widowed", 
"perc_marital_status_male_divorced", "perc_marital_status_female_nevermaried", 
"perc_marital_status_female_maried", "perc_marital_status_female_separated", 
"perc_marital_status_female_widowed", "perc_marital_status_female_divorced"
  ) %>% 
  cor() %>% 
  ggcorrplot(hc.order = TRUE, tl.cex = 15) 

```

```{r edu, fig.height=12, fig.width=12}
XX %>% 
  dplyr::select(
  "perc_enrolled_preschool",
  "perc_enrolled_kindergarden",
  "perc_enrolled_elementary",
  "perc_enrolled_highschool",
  "perc_enrolled_college",
  "perc_edu_9grade",
  "perc_edu_nodiploma",
  "perc_edu_highshool",
  "perc_edu_somecollege",
  "perc_edu_associate",
  "perc_edu_bachelor",
  "perc_edu_gradprofess",
  "perc_edu_highshool_higher",
  "perc_edu_batchelor_higher"
  ) %>% 
  cor() %>% 
  ggcorrplot(hc.order = TRUE, tl.cex = 15) 

```
```{r anchestry, fig.height=12, fig.width=12}
XX %>% 
  dplyr::select(
  "perc_american",
  "perc_arab",
  "perc_czech",
  "perc_danish",
  "perc_dutch",
  "perc_english",
  "perc_french_except_basque",
  "perc_french_canadian",
  "perc_german",
  "perc_greek",
  "perc_hungarian",
  "perc_irish",
  "perc_italian",
  "perc_lithuanian",
  "perc_norwegian",
  "perc_polish",
  "perc_portuguese",
  "perc_russian",
  "perc_scotch_irish",
  "perc_scottish",
  "perc_slovak",
  "perc_subsaharan_african",
  "perc_swedish",
  "perc_swiss",
  "perc_ukrainian",
  "perc_welsh",
  "perc_west_indian_excluding_hispanic_origin_groups"
  ) %>% 
  cor() %>% 
  ggcorrplot(hc.order = TRUE, tl.cex = 15) 

```
```{r anchestry, fig.height=12, fig.width=12}
XX %>% 
  dplyr::select(
"perc_acute_myocardial_infarction",
  "perc_alzheimer_dementia",
  "perc_asthma",
  "perc_atrial_fibrillation",
  "perc_cancer_breast",
  "perc_cancer_colorectal",
  "perc_cancer_lung",
  "perc_cancer_all",
  "perc_ch_obstructive_pulm",
  "perc_chronic_kidney_disease",
  "perc_depression",
  "perc_diabetes",
  "perc_heart_failure",
  "perc_hypertension",
  "perc_ischemic_heart_disease",
  "perc_obesity",
  "perc_osteoporosis",
  "perc_rheumatoid_arthritis",
  "perc_schizophrenia_psychotic_dis",
  "perc_stroke",
  "perc_tobacco_use"
  ) %>% 
  cor() %>% 
  ggcorrplot(hc.order = TRUE, tl.cex = 15) 

```
```{r to_select2}
to_select2 <- c(
  "perc_families",
  "perc_familes_18childreen",
  "perc_married_couples",
  # "perc_married_couples_u18ychildreen",
  "perc_families_only_male",
  # "perc_families_only_male__18ychildreen",
  "perc_families_only_female",
  # "perc_families_only_female__18ychildreen",
  # "perc_non_families_alone",
  # "perc_non_families_alone65y",
  # "perc_non_families_u18y",
  # "perc_non_families_65y",
  # "perc_relationship_householder",
  # "perc_relationship_spouse",
  # "perc_relationship_child",
  # "perc_relationship_other_relatives",
  # "perc_relationship_other_nonrelatives",
  # "perc_relationship_other_unmaried_part",
  # "perc_marital_status_male_nevermaried",
  # "perc_marital_status_male_maried",
  # "perc_marital_status_male_separated",
  # "perc_marital_status_male_widowed",
  # "perc_marital_status_male_divorced",
  # "perc_marital_status_female_nevermaried",
  # "perc_marital_status_female_maried",
  # "perc_marital_status_female_separated",
  # "perc_marital_status_female_widowed",
  # "perc_marital_status_female_divorced",
  "perc_enrolled_preschool",
  "perc_enrolled_kindergarden",
  "perc_enrolled_elementary",
  "perc_enrolled_highschool",
  "perc_enrolled_college",
  # "perc_edu_9grade",
  # "perc_edu_nodiploma",
  # "perc_edu_highshool",
  "perc_edu_somecollege",
  "perc_edu_associate",
  "perc_edu_bachelor",
  # "perc_edu_gradprofess",
  # "perc_edu_highshool_higher",
  # "perc_edu_batchelor_higher",
  "perc_american",
  "perc_arab",
  "perc_czech",
  "perc_danish",
  "perc_dutch",
  "perc_english",
  "perc_french_except_basque",
  "perc_french_canadian",
  "perc_german",
  "perc_greek",
  "perc_hungarian",
  "perc_irish",
  "perc_italian",
  "perc_lithuanian",
  "perc_norwegian",
  "perc_polish",
  "perc_portuguese",
  "perc_russian",
  "perc_scotch_irish",
  "perc_scottish",
  "perc_slovak",
  "perc_subsaharan_african",
  "perc_swedish",
  "perc_swiss",
  "perc_ukrainian",
  "perc_welsh",
  "perc_west_indian_excluding_hispanic_origin_groups",
  # "perc_withcomputer",
  "perc_withinternet",
  # "total_pop",
  # "perc_imm65",
  "total_beds",
  "urban",
  "perc_acute_myocardial_infarction",
  "perc_alzheimer_dementia",
  "perc_asthma",
  "perc_atrial_fibrillation",
  "perc_cancer_breast",
  "perc_cancer_colorectal",
  "perc_cancer_lung",
  "perc_cancer_all",
  "perc_ch_obstructive_pulm",
  "perc_chronic_kidney_disease",
  "perc_depression",
  "perc_diabetes",
  "perc_heart_failure",
  "perc_hypertension",
  "perc_ischemic_heart_disease",
  "perc_obesity",
  "perc_osteoporosis",
  "perc_rheumatoid_arthritis",
  "perc_schizophrenia_psychotic_dis",
  "perc_stroke",
  "perc_tobacco_use",
  # to many NA for these:
  # "urgent_admission",
  # "annual_wellness_visit",
  # "elective_admission",
  # "emergent_admission",
  # "other_admission",
  "perc_pneumococcal_vaccine",
  # "perc_poverty",
  "median_income",
  "pm2.5",
  "summer_temp",
  "summer_hum",
  "winter_temp",
  "winter_hum",
  "age65_over",
  "median_age",
  # "median_age_male", 
  # "median_age_female", 
  "sex_ratio",
  # "age_dependency",                                   
  # "old_age_dependency",                              
  "child_dependency",
  "perc_black",
  "perc_lat",
  "perc_white",
  "total_tests",
  "days_f0"
)
```



```{r corr, warning=FALSE}

M <- XX %>% 
  dplyr::select(!!to_select2) %>% 
  cor(use ="pairwise.complete.obs")

M[upper.tri(M, diag = TRUE)] <- NA

M_tidy <- tidy(M) %>% 
  pivot_longer(cols = -.rownames, values_to = "cor", names_to = "var2") %>% 
  na.omit() %>% 
  rename(var1 = .rownames) %>% 
  filter(var1 != var2) %>% 
  filter(cor >= 0.8 | cor <= -0.8) %>% 
  arrange(desc(var2)) %>% 
  distinct(cor, .keep_all = TRUE) 
M_tidy
```



We have 22 variables that are can potentially eliminated.


# Cast of characters

```{r to_select2_b}
to_select2 <- c(
  "perc_families",
  "perc_familes_18childreen",
  "perc_married_couples",
  # "perc_married_couples_u18ychildreen",
  "perc_families_only_male",
  # "perc_families_only_male__18ychildreen",
  "perc_families_only_female",
  # "perc_families_only_female__18ychildreen",
  # "perc_non_families_alone",
  # "perc_non_families_alone65y",
  # "perc_non_families_u18y",
  # "perc_non_families_65y",
  # "perc_relationship_householder",
  # "perc_relationship_spouse",
  # "perc_relationship_child",
  # "perc_relationship_other_relatives",
  # "perc_relationship_other_nonrelatives",
  # "perc_relationship_other_unmaried_part",
  # "perc_marital_status_male_nevermaried",
  # "perc_marital_status_male_maried",
  # "perc_marital_status_male_separated",
  # "perc_marital_status_male_widowed",
  # "perc_marital_status_male_divorced",
  # "perc_marital_status_female_nevermaried",
  # "perc_marital_status_female_maried",
  # "perc_marital_status_female_separated",
  # "perc_marital_status_female_widowed",
  # "perc_marital_status_female_divorced",
  "perc_enrolled_preschool",
  "perc_enrolled_kindergarden",
  "perc_enrolled_elementary",
  "perc_enrolled_highschool",
  "perc_enrolled_college",
  # "perc_edu_9grade",
  # "perc_edu_nodiploma",
  # "perc_edu_highshool",
  "perc_edu_somecollege",
  "perc_edu_associate",
  "perc_edu_bachelor",
  # "perc_edu_gradprofess",
  # "perc_edu_highshool_higher",
  # "perc_edu_batchelor_higher",
  "perc_american",
  "perc_arab",
  "perc_czech",
  "perc_danish",
  "perc_dutch",
  "perc_english",
  "perc_french_except_basque",
  "perc_french_canadian",
  "perc_german",
  "perc_greek",
  "perc_hungarian",
  "perc_irish",
  "perc_italian",
  "perc_lithuanian",
  "perc_norwegian",
  "perc_polish",
  "perc_portuguese",
  "perc_russian",
  "perc_scotch_irish",
  "perc_scottish",
  "perc_slovak",
  "perc_subsaharan_african",
  "perc_swedish",
  "perc_swiss",
  "perc_ukrainian",
  "perc_welsh",
  "perc_west_indian_excluding_hispanic_origin_groups",
  # "perc_withcomputer",
  "perc_withinternet",
  # "total_pop",
  # "perc_imm65",
  "total_beds",
  "urban",
  "perc_acute_myocardial_infarction",
  "perc_alzheimer_dementia",
  "perc_asthma",
  "perc_atrial_fibrillation",
  "perc_cancer_breast",
  "perc_cancer_colorectal",
  "perc_cancer_lung",
  "perc_cancer_all",
  "perc_ch_obstructive_pulm",
  "perc_chronic_kidney_disease",
  "perc_depression",
  "perc_diabetes",
  "perc_heart_failure",
  "perc_hypertension",
  "perc_ischemic_heart_disease",
  "perc_obesity",
  "perc_osteoporosis",
  "perc_rheumatoid_arthritis",
  "perc_schizophrenia_psychotic_dis",
  "perc_stroke",
  "perc_tobacco_use",
  # to many NA for these:
  # "urgent_admission",
  # "annual_wellness_visit",
  # "elective_admission",
  # "emergent_admission",
  # "other_admission",
  "perc_pneumococcal_vaccine",
  # "perc_poverty",
  "median_income",
  "pm2.5",
  "summer_temp",
  "summer_hum",
  "winter_temp",
  "winter_hum",
  "age65_over",
  "median_age",
  # "median_age_male", 
  # "median_age_female", 
  "sex_ratio",
  # "age_dependency",                                   
  # "old_age_dependency",                              
  "child_dependency",
  "perc_black",
  "perc_lat",
  "perc_white",
  "total_tests",
  "days_f0"
)
```



```{r}
XX <- all_dat4[, names(all_dat4) %in% to_select2] %>% 
  mutate_all(as.numeric)

# x <-  lapply(names(XX), function(x) XX[[x]]) 
# XX <- do.call(rbind, x)

logitZZ <- log(all_dat4$perc_imm65) - log(1-all_dat4$perc_imm65)

ZZ <- all_dat4$perc_imm65

YY <-  all_dat4$deaths

NC <-  all_dat4$cases

NP <-  all_dat4$total_pop
```

# Visualization

Some correlations are computed. Giving the large number of variables is hard to see something. 
** I had to use `na.omit`*

```{r fig.height=12, fig.width=12}
M <- cor(cbind(logitZZ, XX),  use ="pairwise.complete.obs")


ggcorrplot(M, hc.order = TRUE, tl.cex = 7)
```


## adjusted v unadjusted regression analysis

As far as I understand this, the adjust regression includes also other factors than `imm65`. Offset is set to `NC` number of cases or `NP` population.

### Cases

```{r cases_unadjusted}
OutcomeGLM_cases_u = glm( YY ~ ZZ, family = quasipoisson(link="log"), offset = log(NC))
summary(OutcomeGLM_cases_u)
```

```{r cases_adjusted}
OutcomeGLM_cases_a = glm ( YY ~ ZZ + as.matrix(XX), family = quasipoisson(link="log"), offset = log(NC))

tidy(OutcomeGLM_cases_a) %>% 
  mutate(term = stringr::str_remove_all(term, "as\\.matrix\\(XX\\)")) %>% 
  filter(p.value < 0.05) 

# %>% 
#   mutate(degrees = round((atan(estimate) *180 / pi), digit = 0))
```

```{r}
sign_factors <- tidy(OutcomeGLM_cases_a) %>% 
  filter(p.value < 0.05) %>% 
  filter(!term %in% c("ZZ", "(Intercept)")) %>% 
  select(term) %>% 
  mutate(term = stringr::str_remove_all(term, "as\\.matrix\\(XX\\)")) 
```



### Interactions Cases

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(broom)

lin_mod <- function(formula) {
    x <-   glm(formula, offset = log(NC), family = quasipoisson(link="log"))
    tidy(x)
}

 
base_formula <- "YY ~ ZZ + as.matrix(XX)"

all_formulas <- paste0(base_formula, " + ZZ * XX$", unlist(sign_factors))

names(all_formulas) <- all_formulas

all_interactions <- map_dfr(all_formulas, lin_mod, .id = "formula")

all_interactions %>% 
  # filter(str_detect(term, "ZZ")) %>% 
  filter(term == "ZZ") %>% 
  select(formula, term, estimate, p.value) %>% 
  filter(p.value < 0.05) 
```

